Certified Knowledge Compilation \\ with Application to Verified Model Counting

### 2023-03-10 doc/mc-paper.tex
Separate section for Lean

We have developed a program that can generate CRAT descriptions from
the d-DNNF graphs generated by the state-of-the-art knowledge compiler
\dfour, as well as a fast CRAT proof checker.  We evaluate the
capabilities of our tool chain using benchmark formulas from the 2022
standard and weighted model competitions.  Our tools are able to
handle all but the largest d-DNNF graphs generated by \dfour.
Moreover, we 
present a checker that has been formally verified
in the \lean{} proof assistant, thereby creating the
first formally verified toolchain for
% weighted and unweighted
model counting.

@@

\section{A Formally Verified CRAT Checker}

@@

Our ring evaluator can perform both standard and weighted model
counting.  It performs arithmetic over a subset of the rational
numbers we term $\drational$, consisting of numbers of the form $a
\cdot 2^{b} \cdot 5^{c}$, where $a$, $b$, and $c$ are 
integers, and with $a$ implemented to have arbitary range.
Allowing scaling by powers of 2 enables the density computation
and rescaling required for standard model counting.  Allowing scaling
by powers of both 2 and 5 enables exact decimal arithmetic, handling
the weights used in the weighted model counting competitions.

@@

Each node $\nodeu$ in an $SDD$ has children $p_1, p_2, \ldots, p_k$ and
$s_1, s_2, \ldots p_k$, where the Boolean formula $\phi(\nodeu)$ associated
with the node is defined recursively as $\phi(\nodeu) = \bigwedge_{i=1,k}
[\phi(p_i) \land \phi(s_i)]$.  Moreover, each children $p_i$ and $p_j$
such that $1 \leq i < j \leq k$ must satisfy $\modelset(\phi(p_i))
\cap \modelset(\phi(p_j)) = \emptyset$, and for every $p_i$ and $s_j$
such that $1 \leq i,j \leq k$, we must have $\dependencyset(\phi(p_i))
\cap \dependencyset(\phi(s_j)) = \emptyset$.  The POG representation
of such a node could have $k$ product nodes: $t_i = p_i \pand s_i$
for $1 \leq i \leq k$, and then $k-1$ binary sum nodes to form
the sum of $t_1, \ldots, t_k$.

>>>>

We have developed a program that can generate CRAT descriptions from
the d-DNNF graphs generated by the state-of-the-art knowledge compiler
\dfour, as well as a fast CRAT proof checker.  We evaluate the
capabilities of our tool chain using benchmark formulas from the 2022
standard and weighted model competitions.  Our tools are able to
handle all but the largest d-DNNF graphs generated by \dfour.
Moreover, we 
present a CRAT checker, and a POG model counter that have been formally verified
in the \lean{} proof assistant, thereby creating the
first formally verified toolchain for
% weighted and unweighted
model counting.

@@

\input{lean}

@@

Our ring evaluator can perform both standard and weighted model
counting.  It performs arithmetic over a subset of the rational
numbers we term $\drational$, consisting of numbers of the form
$a \cdot 2^{b} \cdot 5^{c}$, for integers $a$, $b$, and $c$, and
with $a$ implemented to have arbitary range.
Allowing scaling by powers of 2 enables the density computation
and rescaling required for standard model counting.  Allowing scaling
by powers of both 2 and 5 enables exact decimal arithmetic, handling
the weights used in the weighted model counting competitions.

@@

%% Each node $\nodeu$ in an $SDD$ has children $p_1, p_2, \ldots, p_k$ and
%% $s_1, s_2, \ldots p_k$, where the Boolean formula $\phi(\nodeu)$ associated
%% with the node is defined recursively as $\phi(\nodeu) = \bigwedge_{i=1,k}
%% [\phi(p_i) \land \phi(s_i)]$.  Moreover, each children $p_i$ and $p_j$
%% such that $1 \leq i < j \leq k$ must satisfy $\modelset(\phi(p_i))
%% \cap \modelset(\phi(p_j)) = \emptyset$, and for every $p_i$ and $s_j$
%% such that $1 \leq i,j \leq k$, we must have $\dependencyset(\phi(p_i))
%% \cap \dependencyset(\phi(s_j)) = \emptyset$.  The POG representation
%% of such a node could have $k$ product nodes: $t_i = p_i \pand s_i$
%% for $1 \leq i \leq k$, and then $k-1$ binary sum nodes to form
%% the sum of $t_1, \ldots, t_k$.
